---
title: 'Exercise 1: Motivating reproducibility'
output:
  html_document:
    fig_height: 3
    fig_width: 6
    highlight: pygments
---

This RMarkdown file contains R code you can use to complete the exercise from earlier. To see the output simply click on **Knit HTML** above. This might prompt you to install and load some required packages, specifically `knitr`. Just click yes, and the a document including this narrative, the R code, and figures should pop up.


## Packages:

We will use the `ggplot2` package to make the plots for these exercises. This is not the only way to make this plot, but the `ggplot2` package has some aesthetic defaults that makes it attractive. Load this package:

```{r message=FALSE}
library(ggplot2)
```

If you'd like to learn more about `ggplot2`, we recommend visiting http://ggplot2.org/, which includes links to documentation about the functions in this package as well as external resources.

We will also use the `dplyr` package for data wrangling, let's load that as well:

```{r message=FALSE}
library(dplyr)
```

Lastly, we will use the `downloader` package for downloading files from a secure URL (https) with a simple commands.

```{r message=FALSE}
library(downloader)
```


## Data: 

First, we download the dataset `gapminder-5060.csv` off the web:
```{r}
download(url = "https://raw.githubusercontent.com/mine-cetinkaya-rundel/reproducibility-uscots15/master/data/gapminder-5060.csv", destfile = "gapminder-5060.csv")
```

The `read.csv` function is used for loading the data. Also note that the name of the file is provided as a character string, hence in quotation marks.

```{r}
gap_5060 = read.csv("gapminder-5060.csv")
```


## Exercise 1: 

**Task: Visualize the relationship between GDP and life expectancy for countries in Europe in 1952 using a scatter plot.**

First, create a subset of `gap_5060` for Europe in 1952 by `filter`ing for rows where `continent == "Europe"` and `year == 1952`, and save the resulting subset as a new data frame called `eu52`:

```{r}
eu_52 <- gap_5060 %>%
  filter(continent == "Europe", year == 1952)
```

Then, create the scatter plot:
```{r}
ggplot(data = eu_52, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()                                          
```

For more on creating scatterplots using the `ggplot2` package, see [here](http://docs.ggplot2.org/current/geom_point.html).


## A brief extension:

**Task: Add year 1967 in another color.**

1. Create a subset of `gap_5060` for Europe in 1952 **and 1967** and call it `eu_5267`. Some hints to get you started:

    * Copy and paste the code from above where you created a subset for the year 1952 only.
    * Update of the logical argument specifying `year == 1952` to instead specify 1952 or 1967: `year %in% c(1952, 1967)`.
    * Update the name of the resulting data frame to `eu_5267`

```{r}

```

2. Create the plot. Some more hints:
    * Copy and paste the code that was used for plotting from above.
    * Update the name of the data frame being referred to in the `ggplot` function.
    * Add another aesthetic to the plot for plotting data from th two years in difference colors: `color = factor(year)`. Note that this says the color of the points should be determined by the year, and that we should consider year as a categorical (factor) variable (either 1952 or 1967).

```{r}
                                        
```

