---
title: 'Exercise 2: Extending your analysis'
output:
  html_document:
    fig_height: 3
    fig_width: 6
    highlight: pygments
---

```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(downloader)
```

## Data: 

Download the necessary files:

```{r}
download(url = "https://raw.githubusercontent.com/mine-cetinkaya-rundel/reproducibility-uscots15/master/data/gapminder-5060.csv", destfile = "gapminder-5060.csv")

download(url = "https://raw.githubusercontent.com/mine-cetinkaya-rundel/reproducibility-uscots15/master/data/gapminder-7080.csv", destfile = "gapminder-7080.csv")

download(url = "https://raw.githubusercontent.com/mine-cetinkaya-rundel/reproducibility-uscots15/master/data/gapminder-90plus.csv", destfile = "gapminder-90plus.csv")
```

and load them using the `read.csv` function:

```{r}
gap_5060 = read.csv("gapminder-5060.csv")
gap_7080 = read.csv("gapminder-7080.csv")
gap_90plus = read.csv("gapminder-90plus.csv")
```

There are a number of similarly named functions for reading files, such as  `read.delim` (tab-delimited files) and `read.table` (allows you to specify the separator), and there are similar solutions for Excel spreadsheets and various other formats.

## Exercise 2

**Task: Append the new data `gapminder-7080.csv` and `gapminder-90plus.csv` to your existing data set.**

The functions `rbind` and `cbind` can be used to combine data frames. `rbind` combines rows (such as when adding now observations to a data set), `cbind` combines by columns (useful when adding new variables to a data set).  

In the current context, we want to add new rows to our data set so we'll use `rbind`. One of the really convenient things about using `rbind` rather than trying to combine the spreadsheets in a program like Excel is that `rbind` matches columns by their names rather than locations, so a simple reordering of columns in one data frame versus another is very simply dealt with.

```{r}
gap_combined <- rbind(gap_5060, gap_7080, gap_90plus)
```

**Task: Create line plots of life expectancy over time for Canada, Mexico, and the United States that run from 1952 to 2007.**

Since we're only interested in data from the US, Canada, and Mexico for the years 1950 through 2007, we first create a list containing names of these countries.

```{r}
country_list <- c('Canada', 'Mexico', 'United States')
```

Then, we subset the combined dataset for these countries.

```{r}
gap_north_america <- gap_combined %>%
  filter(country %in% country_list)
```

Finally, we use the `geom_line()` "geom" from ggplot. "Geom" is short for "geometric object" -- basically something that controls how a plot is drawn. So `geom_line()` draws our data as line plots.

```{r}
ggplot(gap_north_america, aes(x = year, y = lifeExp, color = country)) + 
  geom_line() + 
  xlab("Year") + 
  ylab("Life Expectancy (years)")
```

## Further extension:

**Task: Create the same plot as before, but now also include data from Cambodia, China, Japan, Egypt, South Africa, Uganda.**

Insert the code for creating the same plot as before, but now also including data from Cambodia, China, Japan, Egypt, South Africa, Uganda. Don't forget to save this as a new data frame. Then create the same plot as before. 

Remember that your code must go into a code chunk. You can insert a code chunk by typing the required characters or clicking on *Chunks* and then *Insert Chunk* above.

