---
title: 'Exercise 2: Extending your analysis'
output:
  html_document:
    fig_height: 3
    fig_width: 6
---

Append the new data `gapminder-7080.csv` and and `gapminder-90plus.csv` to your existing data set.

## Reading files

Reading in the data is quick and easy with the `read.csv` function.  
```{r}
gap_5060 <- read.csv('data/gapminder-5060.csv')
gap_7080 <- read.csv('data/gapminder-7080.csv')
gap_90plus <- read.csv('data/gapminder-90plus.csv')
```

There are a number of similarly named functions for reading files, such as  `read.delim` (tab-delimited files) and `read.table` (allows you to specify the separator), and there are similar solutions for Excel spreadsheets and various other formats.

## Combining data frames

The functions `rbind` and `cbind` can be used to combine data frames. `rbind` combines rows (such as when adding now observations to a data set), `cbind` combines by columns (useful when adding new variables to a data set).  

In the current context, we want to add new rows to our data set so we'll use `rbind`. One of the really convenient things about using `rbind` rather than trying to combine the spreadsheets in a program like Excel is that `rbind` matches columns by their names rather than locations, so a simple reordering of columns in one data frame versus another is very simply dealt with.


```{r}
gap_combined <- rbind(gap_5060, gap_7080, gap_90plus)
```

# Task 2

Create line plots of life expectancy over time for Canada, Mexico, and the United States that run from 1952 to 2007.

We're going to using the `ggplot2` library to facilitate efficient figure creation, so let's load the library.
```{r}
library(ggplot2)
```

Since we're only interested in data from the US, Canada, and Mexico for the years 1950 through 2007, we first create a list containing names of these countries.

```{r}
country_list <- c('Canada', 'Mexico', 'United States')
```

Then, we subset the combined dataset for these countries.
```{r}
gap_north_america <- subset(gap_combined, country %in% country_list)
```

Finally, we use the `geom_line()` "geom" from ggplot. "Geom" is short for "geometric object" -- basically something that controls how a plot is drawn. So `geom_line()` draws our data as line plots.

```{r}
ggplot(gap_north_america, aes(x = year, y = lifeExp, color=country)) + 
  geom_line() + 
  xlab("Year") + 
  ylab("Life Expectancy (years)")
```

### Extending life expectancy plots to include more countries

Insert the code for creating the same plot as before, but now also including data from Cambodia, China, Japan, Egypt, South Africa, Uganda. 

Remember that your code must go into a code chunk. You can insert a code chunk by typing the required characters or clicking on Chunks and then Insert Chunk above.

