---
title: "Connecting data, analysis, and results using a reproducible framework"
author: "Mine Çetinkaya-Rundel & Nicholas Horton"
date: "USCOTS 2015"
output:
  pdf_document: default
  html_document:
    highlight: pygments
---

```{r echo=FALSE}
library(knitr)
library(markdown)
library(rmarkdown)
```

```{r echo=FALSE}
rstudio_login <- read.csv("rstudio_login/rstudio_login.csv", stringsAsFactors = FALSE)
url <- rstudio_login[i,1]
pass <- rstudio_login[i,2]
```

#### Materials:

* Slides at http://bit.ly/reproduce_uscots15_slides

* For GitHub users: all source code at http://bit.ly/reproduce_uscots15_git

#### Agenda for hands-on portions of the workshop:

* Exercise 1: Dataset at http://bit.ly/reproduce_uscots15_gap5060

* Revisit Exercise 1: requires RStudio. 
    * You have two options for accessing RStudio:
        * **Option 1: I'm new to R/RStudio:** 
            * Go to `r url`
            * Username: guest
            * Password: `r pass`
            * This account will work until the end of today

        * **Option 2: I already use R/RStudio:** Do you have the **latest** versions of R and RStudio installed locally on your computer?
            * No, or you're not sure $\rightarrow$ Go to Option 1
            * Yes $\rightarrow$ Install and load `knitr`, `ggplot2`, and `downloader` packages
```{r eval = FALSE}
install.packages("knitr") ; library("knitr")
install.packages("ggplot2") ; library("ggplot2")
install.packages("downloader") ; library("downloader")
```   
    * Download RMarkdown template using

```{r eval = FALSE}
download(url = "http://bit.ly/reproduce_uscots15_ex1", destfile = "ex1-template.Rmd")
```    
    

* Exercise 2: Download RMarkdown template using

```{r eval = FALSE}
download(url = "http://bit.ly/reproduce_uscots15_ex2", destfile = "ex2-template.Rmd")
```

#### Questions / Connections:

Feel free to contact Mine at mine@stat.duke.edu after the workshop with any questions.